---
title: "Viz and EDA"
author: Anna Ma
date: Oct 05 2021
output: github_document
---

## GG plot

```{r}
library(tidyverse)
library(ggridges)
```

load a data set that we use the most

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

## Let's make some plots!

### Scaterplot

tmax vs. tmin

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  # tell what varaible to map on the plot 
  geom_point()
```

save ggplot

```{r}
ggp_tmax_tmin = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  # tell what varaible to map on the plot 
  geom_point()

ggp_tmax_tmin
# can use ggsave() here 
```

### LET'S FANCY IT UP

add color? lines?

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  # tell what varaible to map on the plot 
  geom_point(alpha = 0.3) +
  geom_smooth(se = FALSE) + 
  # it's har to see the lines, so we can use other arguments
  facet_grid(.~ name)
  #Using the dot because ...?  
```

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  # tell what varaible to map on the plot 
  geom_point(alpha = 0.3, aes(color = name)) +
  # shifted the aes from global to only the points 
  geom_smooth(se = FALSE)
```

make one more scatter plot

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, size = prcp)) + 
  geom_point(alpha = 0.3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

## use data manipulation as part of this

```{r}
weather_df %>% 
  filter(name == "CentralPark_NY") %>% 
  mutate(
    tmax = tmax * ( 9 / 5 ) + 32,
    tmin = tmin * ( 9 / 5 ) + 32
  ) %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

## Stacking geoms
which geoms do you want 

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_smooth()
```

give u an idea of how dense the data is at a certain point
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_hex()
```

## Univariate plots

```{r}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +
  geom_histogram() +  
  facet_grid(. ~ name)
```

Let's try some other plots

```{r}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +
  geom_density(alpha = 0.3)
```

still with tmax and name

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
```

some people like violin plots???

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmax)) +
  geom_violin()
```


what about ridges ...

```{r}
library(ggridges)
weather_df %>% 
  ggplot(aes(x = tmax, y = name)) +
  geom_density_ridges(alpha = 0.3, scale = .8)
```





