---
title: "2021_10_07_lecture"
subtitle: "Part II"
author: "Anna Ma"
date: "10/10/2021"
output: github_document
---

```{r}
library(tidyverse)
library(ggridges)
library(patchwork)
```

## Load the weather data set

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

weather_df
```

## Plot from last lecture

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5)
```

## Add Labels

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  labs(
    title = "Temperature Plot",
    x = "Minimum daily temperature(C)",
    y = "Maximum daily temperature(C)",
    caption = "Data from rnoaa package; temperature in 2017"
  )
```

## Scales

aesthetic mapping tell you how a variable mapps onto some graph feature. This happens in the background, but we still have some control when we create the graph. Scales is going to control the behavior of things that happen behind the scenes with the sesthetic mappings. 

start with the same plot and define x and y scales

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  labs(
    title = "Temperature Plot",
    x = "Minimum daily temperature(C)",
    y = "Maximum daily temperature(C)",
    caption = "Data from rnoaa package; temperature in 2017"
  ) + 
  # the variable that maps to the x axis is continuous, so use the following 
  scale_x_continuous(
    breaks = c(-15,0,15), # suppose we didn't want -10,0,and 10
    labels = c("-15 C", "0 ","15")
  ) + 
  scale_y_continuous(
    #trans = "sqrt" # using the square root transformation
    position = "right" # put the scale on the right
  )
```

Look at the color scales

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  labs(
    title = "Temperature Plot",
    x = "Minimum daily temperature(C)",
    y = "Maximum daily temperature(C)",
    caption = "Data from rnoaa package; temperature in 2017"
  ) +
  scale_color_hue(
    # we can define something different for the name here
    # we update from the variable "name" to "location"
    name = "Location",
    h = c(100,300)) #h tells something about the hue of a scale, we can define our own color here
```

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  labs(
    title = "Temperature Plot",
    x = "Minimum daily temperature(C)",
    y = "Maximum daily temperature(C)",
    caption = "Data from rnoaa package; temperature in 2017"
  ) +
  # if u don't like what the usual default ggplot color, and not good at define own color and design, this is the package u can use. we can look up the viridis package online 
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE) 
# here, name is a categorical predictor that has three levels, so the scale is discrete, therefore, we have to add discrete = TRUE.
```

## Themes
Plot features that are not data related

Shift the legend

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  labs(
    title = "Temperature Plot",
    x = "Minimum daily temperature(C)",
    y = "Maximum daily temperature(C)",
    caption = "Data from rnoaa package; temperature in 2017"
  ) +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE) +
  theme(legend.position = "bottom")
# there are a lot of other argument to control legend
```

change the overall theme: theme_bw(), theme_minimal(), theme_classic(), ggthemes::

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  labs(
    title = "Temperature Plot",
    x = "Minimum daily temperature(C)",
    y = "Maximum daily temperature(C)",
    caption = "Data from rnoaa package; temperature in 2017"
  ) +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE) +
  theme_minimal() + # this is the overall theme, update at first, if update behind other theme() tweaks, this is going to override everything, if we want to use theme() to change things, put it below the overal theme, for example: 
  theme(legend.position = "bottom")
```

## Setting options

```{r, eval = FALSE}
# put this chunk at the top of the code, then all the plot in the following document will use this theme

library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal()+ theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_color_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Data args in geom()

```{r}
central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

#assume that everything is from waikiki
ggplot(data = waikiki, aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_line()#from left to right connect the dots

#if we want to use one whole data set in geom_point and one whole other data set in geom_line, we can do this:
ggplot(data = waikiki, aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_line(data = central_park)
```

## `patchwork`

refer back to facet, which creating the exact same plot in the different panel, the same aes mapping and geoms to create for different categories. 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = 0.5) +
  facet_grid(. ~ name) # no rows, and have columns according to name
```

if we want to put a density plot aside of a plotting--> when we want multi-panel plot but cannot facet, patchwork

```{r}
tmax_tmin_p =
  weather_df %>% 
    ggplot(aes(x = tmin, y = tmax, color = name)) +
    geom_point(alpha = .5) +
    theme(legend.position = "none")

prcp_dens_p = 
  weather_df %>% 
  filter(prcp > 0) %>% # to get a better graph, we have to know to use the days with precipitation
  ggplot(aes(x = prcp, fill = name)) + #fill the density plot according to name
  geom_density(alpha = .5) +
  theme(legend.position = "none")

tmax_date_p = 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE) + #standard error is false
  theme(legend.position = "none")
```

now we put all those plot together to a multi-panel plot

```{r}
# 3 panel side by side
#tmax_tmin_p + prcp_dens_p + tmax_date_p

#put the last two together
#tmax_tmin_p + (prcp_dens_p + tmax_date_p)

#one on top, next two on bottom
#tmax_tmin_p / (prcp_dens_p + tmax_date_p)

(tmax_tmin_p + prcp_dens_p) / tmax_date_p
```

